Explore consensus cell types

### 2025-01-23 analyses/cell-type-consensus/exploratory-notebooks/02-explore-consensus-results.Rmd
use alphabet palette

# stacked bar chart showing the distribution of the top 9 cell types for each project, including Unknown
celltype_distribution_plot <- function(label,
                                       results_df = all_results_df){

  # filter to get only results for the specified project 
  plot_df <- results_df |> 
    dplyr::filter(project_label == {{label}}) |>
    dplyr::mutate(
      # get most frequently observed cell types across libraries in that project 
      top_celltypes = forcats::fct_lump_n(consensus_annotation, 9, other_level = "All remaining cell types", ties.method = "first") |> 
        # sort by frequency 
        forcats::fct_infreq() |> 
        # make sure all remaining and unknown are last 
        forcats::fct_relevel("All remaining cell types", "Unknown", after = Inf)
    )

  # get color assignments 
  # put grey for unknoqn and all remaining cells at the end of list since they will show up last 
  # first figure out how many cell types are assigned to make sure colors are assigned properly 
  num_assigned_celltypes <- plot_df |> 
    dplyr::filter(!top_celltypes %in% c("Unknown")) |> 
    dplyr::pull(top_celltypes) |> 
    unique() |> 
    length()

  colors <- c(
    palette.colors(palette = "Set1")[1:num_assigned_celltypes], # first get colors for cell types that are not unknown or all remaining
    "grey95" # level for unknown
  ) 
  celltypes <- unique((plot_df$top_celltypes)) |> 
    sort() #make sure they are in the right order 
  names(colors) <- celltypes

@@

```{r}
# list all results files 
results_files <- list.files(opt$results_dir, pattern = "*_consensus-cell-types.tsv.gz", full.names = TRUE)

@@

@@

project_labels |> 
  purrr::map(celltype_distribution_plot)
```

>>>>

# stacked bar chart showing the distribution of the top 9 cell types for each project, including Unknown
celltype_distribution_plot <- function(label,
                                       results_df,
                                       colors){

  # filter to get only results for the specified project 
    plot_df <- plot_df |> 
      dplyr::filter(project_label == {{label}}) |> 
      dplyr::mutate(
        # relevel factors for specific project 
        top_celltypes = forcats::fct_infreq(top_celltypes) |> 
          forcats::fct_relevel("All remaining cell types", "Unknown", after = Inf)
      )

@@

```{r}
# list all results files 
results_files <- list.files(results_dir, pattern = "*_consensus-cell-types.tsv.gz", full.names = TRUE)

@@

```{r}
plot_df <- all_results_df |> 
    dplyr::group_by(project_id) |> 
    dplyr::mutate(
      # get most frequently observed cell types across libraries in that project 
      top_celltypes = forcats::fct_lump_n(consensus_annotation, 9, other_level = "All remaining cell types", ties.method = "first") |> 
        # sort by frequency 
        forcats::fct_infreq() |> 
        # make sure all remaining and unknown are last, use this to assign colors in specific order
        forcats::fct_relevel("All remaining cell types", "Unknown", after = Inf)
    )

# get all unique cell types ordered by frequency 
unique_celltypes <- plot_df |> 
  dplyr::filter(!top_celltypes %in% c("All remaining cell types", "Unknown")) |> 
  dplyr::pull(top_celltypes) |> 
  unique() |>
  sort() |> 
  as.character()

# get color palette
colors <- c(
  palette.colors(palette = "alphabet"),
  "black", # 1 extra since alphabet is 26 and we have 27, this will be plasma cell which shows up once 
  "grey60", 
  "grey95"
)
names(colors) <- c(unique_celltypes, "All remaining cell types", "Unknown")
```

@@

project_labels |> 
  purrr::map(\(label){
    celltype_distribution_plot(label, plot_df, colors)
    })
```
